---
title: "fertility_modelling"
author: "Camille Belmin"
date: "6/7/2021"
output:
  bookdown::pdf_document2:
    toc: false
    latex_engine: xelatex

header-includes: \usepackage{caption} \usepackage{booktabs} \usepackage{xcolor} \usepackage{longtable}
  \usepackage{float} \usepackage{rotating} \floatplacement{figure}{H} \usepackage{lineno}
  \linenumbers
---


input: 
data_model_glm.rds
metadata_dhs_full.csv
age_group_correspondance.csv


output: 
model_glm_simple.rds
model_glm_base.rds
coef_model_glm_simple.rds
model_glm_base_w.rds
model_glm_no_urban.rds
model_glm_sq
model_glm_simple_no_urban.rds
model_glm_alt.rds
model_glm_educ.rds
model_glm_educ_w.rds
model_glm_energy.rds
model_glm_energy_w.rds
model_glm_energy_simple.rds

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(here)
library(miceadds)
library(broom)
library(sjlabelled)
#library(stargazer)
library(imputeTS)

metadata_dhs = read.csv(here("analysis","data", "derived_data","metadata_dhs_full.csv"))
#source.all(here("R"))
source(here("R", "add_cooking_fuel_type.R"))
source(here("R", "recode_na.R"))
path_dhs_data = path_dhs_data = paste0(str_remove(here(), "microsim.energy"), "DHS_data/") # the DHS data is in the parent directory because I use DHS data in several projects

```

# Create the data for the regression from DHS survey

We take the DHS data of Zambia for 7 years and we run a logistic regression to predict the probability of giving birth depending on different covariates: energy access, education and whether the women lives in rural or urban areas.  



```{r cars}


 p_country_code = "ZMB"
 p_min_year = 2001 # ethiopia: 2005. Senegal: 2017
# # metadata of the country 
 p_age_group_map = read.csv(here("analysis", "data", "raw_data", "age_group_correspondance.csv")) 
# 
# p_country_code = "ZMB"
 p_metadata = metadata_dhs
# 
metadata_country = p_metadata %>% filter(country_code == p_country_code, survey_year >= p_min_year)
#   # Read the data from Zambia for the last three waves, and combine them together
   data_pre = NULL
  for( i in 1:nrow(metadata_country)) {
    data_current_raw = readRDS(file =  paste0(path_dhs_data,"IR/", metadata_country$file_name[i])) %>%
      as.data.frame()
    names(data_current_raw) = tolower(names(data_current_raw))
    data_current =  data_current_raw %>%
      dplyr::select(#caseid,
        #contains("v190"),
        year = v007,
        elec = v119, #radio = v120, tv = v121, fridge = v122, #
        contains("v161"),
        #mobile = v169a, internet = v171a, internet_frequency = v171b,
        age = v012, agecat = v013,
        educ_years = v133, birth1 = v209, birth5 = v208, urban = v025, weight1e6 = v005 ) %>%
      {if("v161" %in% names(.))   add_cooking_fuel_type(., coal = "traditional") else .} %>% 
      mutate(survey_id = metadata_country$SurveyId[i])
    print(i)
    data_current = data_current %>% copy_labels(data_pre)
    data_pre = bind_rows(data_pre , data_current)
  }

  # Pre-process the combined data
  data_glm = data_pre %>%
    # make tpye of cooking fuel a binary variables
    mutate(modern_cooking_fuel =
             ifelse(cooking_fuel_type == "modern", 1,
                    ifelse(is.na(cooking_fuel_type), NA, 0))) %>%
    # transform multiple birth into single birth
    mutate(birth1 = ifelse(birth1 %in% c(2,3,4,5,6), 1, birth1),
           urban = ifelse(urban == 2, 0, urban),
           weight = weight1e6/1e6) %>%
    # recode non-numerical variables coded as numbers
    mutate(educ_years = ifelse(educ_years > 90, NA, educ_years)) %>%
    mutate(across(any_of(c("urban", "elec", "radio", "tv", "fridge",
                           "mobile")),
                  recode_na, threshold_NA = 1)) %>%
    mutate(across(any_of(c("elec","agecat",
                           "modern_cooking_fuel", "urban")), as.factor)) %>%
    # mutate(year = as.factor(year)) %>%
    # make education levels. Levels of education in Zambia:
    # 7 complete primary, 9 complete lower sec, 12 complete higher sec
    # according to https://www.scholaro.com/pro/Countries/Zambia/Education-System
    mutate(educ_level = case_when(educ_years == 0 ~ "1",
                                  educ_years > 0 & educ_years < 7 ~ "2",
                                  educ_years == 7 ~ "3",
                                  educ_years >7 & educ_years < 10 ~ "4",
                                  educ_years > 9 & educ_years < 13 ~ "5",
                                  educ_years > 12 ~ "6")) %>%
    # modify the age group variable so that it corresponds to the age groups from 0 to 100 year old
    mutate(age_group_char = case_when(agecat == 1 ~ "15--19",
                                      agecat == 2 ~ "20--24",
                                      agecat == 3  ~ "25--29",
                                      agecat == 4  ~ "30--34",
                                      agecat == 5  ~ "35--39",
                                      agecat == 6  ~ "40--44",
                                      agecat == 7  ~ "45--49")) %>%
    # add the age group as an integer
    left_join(p_age_group_map, by = "age_group_char") %>%
    mutate(age_group = as.factor(age_group))

# Re-weight, so that sum of weight equals sample size, since we pooled samples from different years

sum(data_glm$weight) 
nrow(data_glm)

data_glm %>% group_by(survey_id) %>% summarize(sum_w = sum(weight))
 

data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)


```

# Regressions 

## Model without interaction terms

```{r}

# Model without interaction terms with age and energy
model_glm_simple <- glm(birth1 ~ age_group + educ_level + elec + modern_cooking_fuel + urban + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)

summary(model_glm_simple)

### throw out all the stuff we dont need
sg = model_glm_simple

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_simple.rds"))


```

## Model with interaction terms and URBAN

```{r}

data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)

# Model without interaction terms with age and energy
model_glm_base <- glm(birth1 ~ age_group + educ_level + elec*age_group + modern_cooking_fuel*age_group + urban + year, 
                     family="binomial",
                     data=data)

summary(model_glm_base)

### throw out all the stuff we dont need
sg = model_glm_base

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_base.rds"))

# Same model but weighted regression 
model_glm_base_w <- glm(birth1 ~ age_group + educ_level + elec*age_group + modern_cooking_fuel*age_group + urban + year, 
                     family="binomial",
                     data=data, 
                     weights = weight )
summary(model_glm_base_w)
### throw out all the stuff we dont need
sg = model_glm_base_w

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_base_w.rds"))

summary(model_glm_base_w)
summary(model_glm_base)

# compare coeff using 
# # extract the coefficients only and make it a list

tidy_model_glm_coef = tidy(model_glm_base) %>%
  select(term,estimate_base = estimate) %>% 
  left_join(tidy(model_glm_base_w) %>%
  select(term,estimate_base_w = estimate), by = "term") %>% 
  pivot_longer(cols = -term, names_to = "model", values_to = "estimate")

tidy_model_glm = tidy(model_glm_base) %>%
  # retrieve the std errors
  select(term, std_error_base = std.error) %>% 
  left_join(tidy(model_glm_base_w) %>%
  select(term, std_error_base_w = std.error), by = "term") %>% 
  pivot_longer(cols = -term, names_to = "model", values_to = "std_error") %>% 
  add_column( estimate = tidy_model_glm_coef$estimate) %>% # put the estimate
  mutate(model = str_remove(model, "std_error_")) # rename


ggplot(data = tidy_model_glm %>% filter(!(term == "age_group10:modern_cooking_fuel1")), aes(x = term, y = estimate, color = model)) +
  geom_point(alpha = 0.5)+
  geom_errorbar(aes(ymin = estimate - 0.95*std_error, ymax = estimate + 0.95*std_error ))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Zoom in 
ggplot(data = tidy_model_glm %>% filter(!(term == "age_group10:modern_cooking_fuel1")), aes(x = term, y = estimate, color = model)) +
  geom_point(alpha = 0.5)+
  geom_errorbar(aes(ymin = estimate - 0.95*std_error, ymax = estimate + 0.95*std_error ))+
  ylim(-2.5,2.5)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# CONCLUSION: not much difference in the coefficients between 

```

## Model with interaction terms and NO urban

```{r}
data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)

# Model WITH interaction terms but without urban
model_glm_no_urban <- glm(birth1 ~ age_group + educ_level + elec*age_group + modern_cooking_fuel*age_group + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)

summary(model_glm_no_urban)

### throw out all the stuff we dont need
sg = model_glm_no_urban

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_no_urban.rds"))

```

## Model without interaction terms and NO urban

```{r}
# Model without urban as a predictor
model_glm_simple_no_urban <- glm(birth1 ~ age_group + educ_level + elec + modern_cooking_fuel + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data_glm)

summary(model_glm_simple_no_urban)

### throw out all the stuff we dont need
sg = model_glm_simple_no_urban

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_simple_no_urban.rds"))
```

## Model with interaction terms, urban and square of year

```{r}

# Model with square opf year
model_glm_sq <- glm(birth1 ~ age_group + educ_level + elec*age_group + modern_cooking_fuel*age_group + urban + year_sq, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)

summary(model_glm_sq)

### throw out all the stuff we dont need
sg = model_glm_sq

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_sq"))

```

```{r}
# Alternative model with only the last year of DHS
model_glm_alt <- glm(birth1 ~ age_group + educ_level + elec*age_group + modern_cooking_fuel*age_group + urban + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data %>% filter(year %in% c(2018,2019)))

summary(model_glm_alt)

### throw out all the stuff we dont need
sg = model_glm_alt

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

rm(list=ls(envir = attr(sg$terms, ".Environment")),
      envir = attr(sg$terms,
               ".Environment"))

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_alt.rds"))

``` 

## Model without energy as predictor, without urban

```{r}
# Model without energy as predictor (so only age and education). 
# used to compare our results with those from WIC
data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)

model_glm_educ <- glm(birth1 ~ age_group + educ_level + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)


summary(model_glm_educ)

# Throw out all the stuff we don't need
sg = model_glm_educ

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

rm(list=ls(envir = attr(sg$terms, ".Environment")),
      envir = attr(sg$terms,
               ".Environment"))

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_educ.rds"))


# Same, with urban and with weights 

model_glm_educ_w <- glm(birth1 ~ age_group + educ_level + year + urban, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data, weights = weight)


summary(model_glm_educ_w)

# Throw out all the stuff we don't need
sg = model_glm_educ_w

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL


write_rds(sg, here("analysis", "data", "derived_data", "model_glm_educ_w.rds"))


```

## Model without education as a predictor, and interaction terms, No urban

```{r}
# Model without education as predictor (so only age and energy). 
# used to compare our results with those from WIC
data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)

model_glm_energy <- glm(birth1 ~ age_group + elec*age_group + modern_cooking_fuel*age_group + year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)


summary(model_glm_energy)

# Throw out all the stuff we don't need
sg = model_glm_energy

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_energy.rds"))

# Same but with weights and with urban

model_glm_energy_w <- glm(birth1 ~ age_group + elec*age_group + modern_cooking_fuel*age_group + year + urban, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data, weights = weight)


summary(model_glm_energy_w)

# Throw out all the stuff we don't need
sg = model_glm_energy_w

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_energy_w.rds"))

```

## Model without education as a predictor, WITHOUT interaction terms, No urban

```{r}
# Model without education as predictor (so only age and energy). 
# used to compare our results with those from WIC
data = readRDS(here("analysis", "data", "derived_data", "data_model_glm.rds")) %>% na.omit() %>% 
  mutate(year_sq = year*year)

model_glm_energy_simple <- glm(birth1 ~ age_group + elec + modern_cooking_fuel+  year, #+ year ,#+ year*agecat,
                     family="binomial",
                     data=data)


summary(model_glm_energy_simple)

# Throw out all the stuff we don't need
sg = model_glm_energy_simple

sg$residuals = NULL
sg$weights = NULL
sg$fitted.values = NULL
sg$prior.weights = NULL
sg$na.action = NULL
sg$linear.predictors = NULL
sg$fitted.values = NULL
sg$effects = NULL
sg$data = NULL
sg$qr$qr = NULL
sg$y = NULL

write_rds(sg, here("analysis", "data", "derived_data", "model_glm_energy_simple.rds"))


```


